---
title: "Capstone Exercise"


knit: (function(input, ...){
  rmarkdown::render(input, output_format = "bookdown::word_document2", output_options = list(number_sections = FALSE), output_file = "Capstone.docx"); doconv::docx2pdf("Capstone.docx")
  })
---

# Capstone Exercise

The final exercise for this course involves performing a time series analysis on real-world data: Carbon Dioxide concentration at the Mauna Loa Observatory, from early 1959 to Present. You'll go through the process of importing the data, converting to a time series object (with `zoo`), imputing missing values, and plotting the resulting data. Additionally, you will create an aggregate of the data, as well as a rolling window average of the data. 

## Importing the Data

```{r, echo = TRUE, message=F, warning=F}
# The following libraries are included for you

library(zoo)
# Sample data from the Mauna Loa Observatory
# https://gml.noaa.gov/webdata/ccgg/trends/co2/co2_mm_mlo.csv

# Data is already pre-processed as a `zoo` object. It contains missing values,
# so we'll need to impute those!

# This will be hidden from the users, of course.
missing_co2 <- readRDS("data/missing.Rds")
```

\newpage

## Visualizing the Data

Perform basic data exploration by:

  1. Printing the first 20 observations with `head()`, and,
  2. Creating a plot of the data with `plot()`
  
```{r q1-1}
head(missing_co2, n = 20)
```

```{r q1-2}
plot(missing_co2)
```

Notice that there are "holes" in the data? this suggests that we've got NA values, which is apparent when we view the first few observations with `head()`.

\newpage

## Imputing the Missing Values

Impute the missing values with a *cubic spline* interpolation, then plot the results

```{r q2}
filled_co2 <- na.spline(missing_co2)
```

```{r q2-2}
plot(filled_co2)
```

\newpage

## Yearly Aggregate

Using `aggregate()`, create a yearly mean of the data, then plot the data

*Hint: use `lubridate::year()`

```{r q3-1}
yearly_co2 <- aggregate(filled_co2,
  by = lubridate::year,
  FUN = mean
)


plot(yearly_co2)
```

\newpage

## Rolling Window

Calculate a 12-month rolling window average of the data, then overlay the results (in blue) on top of the original data. Label the x-axis as "Time" and the y-axis as "CO2 Concentration", and give your graph the title "Monthly CO2 Concentration at Mauna Loa Observatory"

```{r q4}
roll <- rollapplyr(filled_co2,
  FUN = mean,
  width = 12
)

plot(filled_co2,
     xlab = "Time",
     ylab = "CO2 Concentration",
     main = "CO2 Concentration at Mauna Loa Observatory")

lines(roll, col = "blue")

```

\newpage

## Given Code

The following code is given to the learners at the beginning of the exercise:

```{r eval = FALSE}

# Question 1: Explore the Data
___(missing_co2, ___ = ___)

___(missing_co2)

```

```{r eval = FALSE}
# Question 2: Impute Missing Values

filled_co2 <- ___(___)

___(filled_co2)
```

```{r eval = FALSE}
# Question 3: Find Yearly Mean Aggregate
yearly_co2 = aggregate(___,
                       by = ___,
                       FUN = ___)

___(___)
```

```{r eval = FALSE}
# Question 4: Calculate a Rolling Window
roll <- rollapplyr(filled_co2,
                   FUN = ___,
                   width = 12)

plot(___,
     xlab = ___,
     ylab = ___,
     main = "CO2 Concentration at Mauna Loa Observatory")

lines(___, col = ___)

```

